Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-08-09 GWLDCII.tex
things in section 5 changed as we were talking over zoom

{\em Proof of claim}: By the structure of determinants we know that $\Delta_{I_i} = fg$, where $g$ involves only variables associated to propagators not in $\mathcal{P}'$ and associated to vertices not in $J$.  \todo{Is the heart of this proof due to the fact that there are no cancelations? It feels tautological to me, but I'm clearly not following}

@@

All that remains is to prove that $\Delta_{I_i}$ has no irreducible factors of degree 3 or more.  Suppose for a contradiction that $f$ is a factor of $\Delta_{I_i}$ of degree $\geq 3$. Note that by removing the propagators which come before those contributing to $f$ in the order given by the construction of $I_i$ and \hlfix{changing $i$ to be the first vertex}{does this generalize to other $i$?} which contributes to $f$, we obtain a \hlfix{different admissible diagram}{Is this just the diagram gotten by removing certain props?} for which $f$ still divides $\Delta_{I_i}$ but also $i\in I_i$ and
$i$ contributes to $f$. \note{The point of all the stuff you were objecting to is just to spell out ``WLOG we may assume $i\in I_i$ and $i$ contributes to $f$''.  How do you want that said?} Showing that this different admissible diagram gives a contradiction is sufficient, and so we may assume that $i\in I_i$ and $i$ contributes to $f$.  Finally, we can suppose that $W$ is minimal in number of propagators with an irreducible factor of degree $\geq 3$.

@@

\textbf{Case 2}: Suppose $p$ has one end on the edge $(i, i+1)$.  If no other propagators are supported on $i$ then the column of $C(W)$ corresponding to vertex $i$ has only one nonzero entry in it, and so $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$; as above, this is a contradiction.  Thus there is at least one other propagator on edge $(i,i+1)$ since since $I_i(p)=i$, one of these other propagators contributes $i+1$ in $I_i$, sob we can take $q$ to be the propagator such that $I_i(q)=i+1$. {\color{violet} How do we know this exists? Why can't we solve this by having $r$ supported on the $i-1^{th}$ edge?}\note{I tried to clarify what $q$ is and why it exists, but I don't know what $r$ is in your question.} We know that $q$ has one end on the edge $(i, i+1)$ and is adjacent to $p$ on that edge in the counterclockwise direction (see Figure \ref{fig no big factors}).  Write $(i, i+1, j, j+1)$ for the support of $q$.  The situation for $q$ is very similar to case 1: in particular, we have $I_{i+1}(q) = j$ by Lemma~\ref{vertex cyclic int lem} and so if $j\in I_i$ then the propagator which contributes $j$ is outside of $q$.

Similarly to Case 1, let \st{$S$ be the set of propagators inside $q$ along with $p$ and $q$ themselves.}{\color{violet}$S = \cP_{in}(q)$ in the $<_i$ order. Note that $p \in S$.}\note{No, $S=\cP_{in}(q)\cup \{p\}$ in the $<_i$ order because $q$ is not inside $p$ in that order}\hlfix{Then all propagators in $S$ are mapped by $I_i(\cdot)$ to vertices strictly before $j$ and no other propagators are supported on vertices strictly before $j$.}{???} \note{This is the usual argument...} Thus the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the submatrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger (with respect to $>_i$) belongs to $I_i$ then $B$ and $C$ are at least one column wide, and so the block form of the matrix gives a nontrivial factorization of $\Delta_{I_i}$.  This yields a contradiction as in Case 1: either $W$ contains unnecessary propagators which contradicts our minimality assumption, or $\det A$ is quadratic which contradicts the assumption that $p$ and $i$ contribute to $f$, an irreducible factor of degree at least 3.

@@

\item Suppose propagators $p$ and $q$ are supported on a common edge $(a,b)$, with $I_i(p)=a$ and $I_i(q)=b$.  Let $x_{p,a},x_{p,b},x_{q,a},x_{q,b}$ be the associated variables in $C(W)$. For any fixed bijection $\sigma$ from $\cP-\{p,q\}$ to $I_i -\{a,b\}$ for which each propagator is supported on its image under the bijection, we can extend $\sigma$ to a bijection of all propagators with $I_i$ in two ways: either $p\mapsto a$ and $q\mapsto b$ or $p\mapsto b$ and $q\mapsto a$.  The sum of the contributions of all these bijections to $\Delta_{I_i}$ is therefore the product of \hlfix{$x_{p,a}x_{q,b}-x_{p,b}x_{q,a}$}{why? terms in determinant come from bijections. Why should all bijections appear in det?}\note{All terms appear by the permuation form of the determinant -- am I mising what you're asking?} with the minor coming from $\cP-\{p, q\}$ and $I_i - \{a,b\}$.  Since there is no cancellation of terms in the expansion of $\Delta_{I_i}$, if any other terms appear then they will cause a factor which is not in the form described in the previous part.  Therefore no such terms exist and $x_{p,a}x_{q,b}-x_{p,b}x_{q,a}$ is a factor of $\Delta_{I_i}$.

@@

Finally then we need to check the scalar.  By Definition~\ref{def R(W)} each linear factor appears with coefficient $1$ and each $2\times 2$ deteriminant factor appears with the same sign as the determinant of the corresponding minor in $C(W)$.  Therefore $R=R(W)$. \todo{I'm pretty sure that there isn't a sign issue here. But do we have a proof?}\note{The previous sentence is the proof -- again I think I'm missing something in your question}
\end{proof}

>>>>

{\em Proof of claim}: By the structure of determinants we know that $\Delta_{I_i} = fg$, where $g$ involves only variables associated to propagators not in $\mathcal{P}'$ and associated to vertices not in $J$.

@@

All that remains is to prove that $\Delta_{I_i}$ has no irreducible factors of degree 3 or more.  Suppose for a contradiction that $f$ is a factor of $\Delta_{I_i}$ of degree $\geq 3$. Without loss of generality\footnote{This is without loss of generality as by removing the propagators which come before those contributing to $f$ in the order given by the construction of $I_i$ and changing $i$ to be the first vertex which contributes to $f$, we obtain an admissible diagram for which $f$ still divides $\Delta_{I_i}$ but also $i\in I_i$ and
$i$ contributes to $f$, and this diagram can be used in place of $W$.} we may assume $i\in I_i$ and $i$ contributes to $f$.
Finally, we can suppose that $W$ is minimal in number of propagators with an irreducible factor of degree $\geq 3$.

@@

\textbf{Case 2}: Suppose $p$ has one end on the edge $(i, i+1)$.  If no other propagators are supported on $i$ then the column of $C(W)$ corresponding to vertex $i$ has only one nonzero entry in it, and so $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$; as above, this is a contradiction.  Since $I_i(p) = i$ there can be no propagator supported on $i$ with an end on the edge $(i-1, i)$.  Thus to have another propagator supported on $i$ there is at least one other propagator on edge $(i,i+1)$.  Further since since $I_i(p)=i$, one of these other propagators contributes $i+1$ in $I_i$, so we can take $q$ to be the propagator such that $I_i(q)=i+1$.  We know that $q$ has one end on the edge $(i, i+1)$ and is adjacent to $p$ on that edge in the counterclockwise direction (see Figure \ref{fig no big factors}).  Write $(i, i+1, j, j+1)$ for the support of $q$.  The situation for $q$ is very similar to case 1: in particular, we have $I_{i+1}(q) = j$ by Lemma~\ref{vertex cyclic int lem} and so if $j\in I_i$ then the propagator which contributes $j$ is outside of $q$.

Similarly to Case 1, let $S = \cP_{in}(q) \cup \{p\}$ in the $<_i$ order. Then all propagators in $S$ are mapped by $I_i(\cdot)$ to vertices strictly before $j$ and no other propagators are supported on vertices strictly before $j$.  Thus the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the submatrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger (with respect to $>_i$) belongs to $I_i$ then $B$ and $C$ are at least one column wide, and so the block form of the matrix gives a nontrivial factorization of $\Delta_{I_i}$.  This yields a contradiction as in Case 1: either $W$ contains unnecessary propagators which contradicts our minimality assumption, or $\det A$ is quadratic which contradicts the assumption that $p$ and $i$ contribute to $f$, an irreducible factor of degree at least 3.

@@

\item Suppose propagators $p$ and $q$ are supported on a common edge $(a,b)$, with $I_i(p)=a$ and $I_i(q)=b$.  Let $x_{p,a},x_{p,b},x_{q,a},x_{q,b}$ be the associated variables in $C(W)$. For any fixed bijection $\sigma$ from $\cP-\{p,q\}$ to $I_i -\{a,b\}$ for which each propagator is supported on its image under the bijection, we can extend $\sigma$ to a bijection of all propagators with $I_i$ in two ways: either $p\mapsto a$ and $q\mapsto b$ or $p\mapsto b$ and $q\mapsto a$.  The sum of the contributions of all these bijections to $\Delta_{I_i}$ is therefore the product of $x_{p,a}x_{q,b}-x_{p,b}x_{q,a}$ \note{Could this use a ref back to where we discuss this kind of arugment in more detail.} with the minor coming from $\cP-\{p, q\}$ and $I_i - \{a,b\}$.  Since there is no cancellation of terms in the expansion of $\Delta_{I_i}$, if any other terms appear then they will cause a factor which is not in the form described in the previous part.  Therefore no such terms exist and $x_{p,a}x_{q,b}-x_{p,b}x_{q,a}$ is a factor of $\Delta_{I_i}$.

@@

Finally then we need to check the scalar.  By Definition~\ref{def R(W)} each linear factor appears with coefficient $1$ and each $2\times 2$ determinant factor appears with the same sign as the determinant of the corresponding minor in $C(W)$.  Therefore $R=R(W)$.
\end{proof}
